---
title: "Markdown report"
output: github_document
---
Figures and tables for the Fenton-adsorption analysis

```{r knitrsetup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE,
                      collapse = TRUE,
                      comment = "#>",
                      fig.path = "images/markdown/"
                      )

# library(wesanderson)
# devtools::install_github("kaizadp/soilpalettes")
```

```{r input}
source("code/0-packages.R")
library(soilpalettes)
theme_set(theme_bw())

# INPUT
meta_elements = read.csv(ELEMENTS)
meta = read.csv(FTICR_META)
hcoc = read.csv(HCOC)
master = read.csv(FTICR_MASTER_LONG)# <- "fticr/fticr_master_long.csv" 
class = read.csv(CLASS)# <- "fticr/meta_class.csv"

relabund = read.csv(RELATIVE_ABUND)# <- "fticr/fticr_relative_abundance.csv"
elements = read.csv(SUMMARY_ELEMENTS)# <- "fticr/fticr_summary_elements.csv"

master_hcoc = read.csv(MASTER_HCOC) 
fenton_hcoc = master_hcoc %>% 
  filter(Treatment %in% c("PreFenton","PostFenton")) %>%
  mutate(Treatment=factor(Treatment, levels = c("PreFenton","PostFenton")))
fentonloss = read.csv(FENTON_LOSS)
sorbed = read.csv(GOETHITE_ADSORPTION)
sorbed_relabund = read.csv(GOETHITE_RELABUND)
```

---

# DOC concentrations
```{r}
source("code/3-doc_conc.R")

print("concentrations, mg/L")
conc_summary %>% 
  dplyr::mutate(wsoc_mg_L = paste(round(doc_mg_L,0),"\U00B1",round(doc_se,1))) %>% 
  dplyr::select(Forest, Fenton, wsoc_mg_L) %>% 
  knitr::kable(align = "r")

print("adsorbed C, mg/g goethite")
conc_summary %>% 
  dplyr::mutate(adsorbed = paste(round(adsorbed_mgg,1),"\U00B1",round(adsorbed_se,1))) %>% 
  dplyr::select(Forest, Fenton, adsorbed) %>% 
  knitr::kable(align = "r")  
```

---

# FTICR-MS
## FTICR spectra

```{r fticr_spectra}
library(scales)
rawmasterlong = read.csv(FTICR_RAWMASTER_LONG) %>% 
  group_by(Forest, Treatment, Fenton, Goethite, Mass) %>% 
  dplyr::summarise(intensity = mean(intensity)) %>% 
  ungroup %>% 
  mutate(Fenton = factor(Fenton, levels = c("PreFenton", "PostFenton")),
         Goethite = factor(Goethite, levels = c("PreGoethite", "PostGoethite")))

rawmasterlong %>% 
  ggplot(aes(x = Mass, xend = Mass, y = 0, yend = intensity/1000000))+
  geom_segment(alpha = 0.5, size = 0.5)+
  scale_color_manual(values = soil_palette("podzol",2))+

  scale_y_continuous(limits = c(0,100), oob = rescale_none, label = comma)+
  labs(y = expression("intensity (x 10" ^6 *")"),
       x = "m/z")+
  facet_grid(Goethite ~ Fenton)+
  theme_kp()+
  NULL
```

## Summary tables
```{r summarytable}
print("average element occurrence")

elements %>% 
  dplyr::mutate(sp = paste(Forest, Treatment),
                val = paste(avg)) %>% 
  select(element, sp, val) %>% 
  dplyr::mutate(sp = factor(sp, levels = c("HW PreFenton","HW PostFenton", "SW PreFenton", "SW PostFenton"))) %>% 
  spread(sp, val) %>% 
  knitr::kable(align = "r")

print("average relative abundance")
relabund %>% 
  ungroup %>% 
  dplyr::mutate(sp = paste(Forest, Fenton),
                val = paste(rel_abund)) %>% 
  select(Class,sp, val) %>% 
  dplyr::mutate(sp = factor(sp, levels = c("HW PreFenton","HW PostFenton", "SW PreFenton", "SW PostFenton")), 
                  Class = factor(Class, levels = c("Condensed Ar",
                                                 "Aromatic",
                                                 "Lignin-like",
                                                 "Carbohydrate-like",
                                                 "Aliphatic-noN",
                                                 "Aliphatic+N"))) %>% 
  spread(sp, val) %>% 
  knitr::kable(align = "r")

print("number of peaks")
master %>% 
  mutate(Treatment = factor(Treatment, levels = c("PreFenton", "PostFenton", "PreFentonGoethite", "PostFentonGoethite"))) %>% 
  group_by(Forest, Treatment) %>% 
  dplyr::summarise(peaks = n()) %>% 
  spread(Forest, peaks) %>% 
  knitr::kable(align = "r")
```

## Van Krevelen domains
1. classes

```{r vk_domains_classes}
meta %>% 
  filter(!is.na(Class)) %>% 
  gg_vankrev(aes(x = OC, y = HC, color = Class))+
  scale_color_manual(values = PNWColors::pnw_palette("Sailboat",6))+
  theme_kp()+
  theme(legend.position = "right")+
  NULL
```

2. DBE

```{r vk_domains_dbe, fig.height=5, fig.width=6}
gg_vankrev(meta, aes(x = OC, y = HC, color = (DBE)))+
  scale_color_gradientn(colors = PNWColors::pnw_palette("Sunset2",6))

```

3. NOSC

```{r vk_domains_nosc, fig.height=5, fig.width=6}
gg_vankrev(meta, aes(x = OC, y = HC, color = NOSC))

```


## Effect of Fenton oxidation 
### Pre- vs. post-Fenton Van Krevelen plots
```{r native_van_krev, fig.cap="native SOM", fig.width=10, fig.height=5, include=FALSE, eval=FALSE}
gg_vankrev(fenton_hcoc[fenton_hcoc$Treatment=="PreFenton",], aes(x=OC,y=HC, color = Treatment))+
  facet_wrap(~Forest)+
  scale_color_manual(values = "dodgerblue1")+
  labs(title = "Pre-Fenton peaks")+
  #guides(colour = guide_legend(override.aes = list(alpha = 1)))+
  theme_kp()+
  theme(legend.position = "none")
```

```{r fenton_van_krev, fig.cap="fenton SOM", fig.width=10, fig.height=5, include=FALSE, eval=FALSE}
gg_vankrev(fenton_hcoc[fenton_hcoc$Treatment=="PostFenton",], aes(x=OC,y=HC, color = Treatment))+
  facet_wrap(~Forest)+
  scale_color_manual(values = "tomato")+
  labs(title = "Post-Fenton peaks")+
  theme_kp()+
  theme(legend.position = "none")
```

```{r vk_fenton_trt, fig.width=9, fig.height=5}
percentile = read.csv(PERCENTILE) %>% mutate(Treatment = factor(Treatment, levels = c("PreFenton", "PostFenton")))
gg_vankrev(percentile, aes(x=OC,y=HC, color = Forest))+
  facet_grid(.~Treatment)+
  stat_ellipse(level = 0.95)+
  #scale_color_manual(values = c("dodgerblue1","tomato"))+
  scale_color_manual(values = soil_palette("podzol",2))+
  guides(colour = guide_legend(override.aes = list(alpha = 1)))+
  theme_kp()+
  
  annotate("text", label = "aliphatic", x = 1.1, y = 2.0)+
  annotate("text", label = "lignin-like", x = 1.1, y = 1.0)+
  annotate("text", label = "aromatic", x = 1.1, y = 0.53)+
  annotate("text", label = "condensed \n aromatic", x = 1.1, y = 0.2)
```

### Fenton loss/gain 

1. fenton loss-gain

```{r fenton_loss, fig.width=5, fig.height=5, include=FALSE}

fentonloss2 = fentonloss %>% 
  group_by(Mass, loss) %>% 
  dplyr::summarize(n = n()) %>% 
  left_join(select(meta_elements, Mass,O), by = "Mass") %>% 
  left_join(hcoc, by = "Mass")

gg_loss =
  gg_vankrev(fentonloss2[fentonloss2$loss=="lost"|fentonloss2$loss=="gained",], 
           aes(x = OC, y = HC, color = loss))+
  #scale_shape_manual(values = c(16,1))+
  #scale_color_brewer(palette = "Dark2")+
  stat_ellipse(level = 0.95)+
  scale_color_manual(values = soil_palette("rendoll",2))+

  theme_kp()+
  theme(legend.position = c(0.7, 0.92),
        legend.direction = "horizontal")+
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

gg_loss2 = 
  ggMarginal(gg_loss,groupColour = TRUE,groupFill = TRUE, size = 10)
```

```{r fenton_oxygen2, fig.width=5, fig.height=5, include=FALSE}

fentonloss2 %>% 
  group_by(loss,O) %>% 
  dplyr::summarize(counts = n()) %>% 
  dplyr::mutate(counts = as.numeric(counts))%>% 
  ungroup() %>% 
  group_by(loss) %>% 
  dplyr::mutate(cumsum = cumsum(counts),
                counts2 = if_else(loss == "gained",counts,-1*counts))->
  loss_summary2

gg_loss_o =
  ggplot(loss_summary2[loss_summary2$loss=="lost"|loss_summary2$loss=="gained",], 
       aes(x=as.factor(O), y=counts2, fill = loss))+
 geom_segment(x = 13.5, y = 0, xend = 13.5, yend = 140,color="black",linetype="longdash")+
  geom_segment(x = 8.5, y = 0, xend = 8.5, yend = -190,color="black",linetype="longdash")+
  annotate("text", label = "median 13", x = 13.5, y = 150)+ 
  annotate("text", label = "median 8", x = 8.5, y = -200)+
  
  geom_bar(stat = "identity", alpha=0.8)+
  #scale_fill_brewer(palette="Dark2")+
  scale_fill_manual(values = soil_palette("rendoll",2))+

  scale_x_discrete(breaks = c("0", "5","10","15","20","25","30"))+
  ylab("number of peaks")+
  xlab("number of O")+
  
  theme_kp()+
  theme(legend.position = c(0.8, 0.3))+
  guides(colour = guide_legend(override.aes = list(alpha = 1)))


```

```{r fenton_loss_combined, fig.width=9, fig.height=4}
# library(patchwork)
# gg_loss2 + gg_loss_o

library(cowplot)
plot_grid(gg_loss2, gg_loss_o, labels = c('A', 'B'), label_size = 12, axis = "bl", align = "hv")

```

2. fenton peaks lost -- combined for HW and SW

```{r fenton_lostpeaks}
fentonloss %>% 
  group_by(Forest, Mass,loss) %>% 
  dplyr::summarise(n = n()) %>% 
  na_if(.,"") %>% 
  na.omit() %>% 
  ungroup() %>% 
  left_join(class, by = "Mass") %>% 
  distinct(Mass, loss, Class) %>% 
  group_by(loss, Class) %>% 
  dplyr::summarize(count = n()) %>% 
  spread(loss, count) %>% 
  knitr::kable(align = "r")


```

```{r fenton_2, eval=FALSE}
## 2. fentons-oxygen

fentonloss2 %>% 
  left_join(class, by = "Mass")-> fenton_el2

ggplot(fenton_el2, aes(x = O, fill = loss))+
  geom_histogram()+
  facet_grid(loss~Class)


master %>% 
  left_join(class, by = "Mass") %>% 
  left_join(meta_elements, by = "Mass")->master2

ggplot(master2, aes(x = O))+
  geom_histogram()+
  facet_grid(Treatment~Class)

```

```{r, eval=FALSE}
## 3. fenton vs. oxygens by class

ggplot(master2[master2$Treatment=="PreFenton"|master2$Treatment=="PostFenton",], 
       aes(x = O, color = Treatment, fill = Treatment))+
  geom_histogram(binwidth = 1, alpha = 0.2, position = "identity")+
  facet_wrap(~Class)

master2 %>% 
  group_by(Mass,Treatment, Class) %>% 
  dplyr::summarise(O = mean(O)) %>% 
  ungroup %>% 
  group_by(Treatment, Class) %>% 
  dplyr::summarise(medianO = median(O)) %>% 
  spread(Treatment, medianO)
```


### NOSC

```{r fenton_nosc_files}
fenton_loss = read.csv(FENTON_LOSS)

loss = fenton_loss %>% 
  filter(loss %in% c("lost", "gained")) %>% 
  left_join(dplyr::select(meta, Mass, NOSC, Class),  by = "Mass")

master_hcoc = read.csv(MASTER_HCOC)
fenton = 
  master_hcoc %>% 
  filter(Treatment %in% c("PreFenton", "PostFenton")) %>% 
  left_join(dplyr::select(meta, Mass, NOSC, Class),  by = "Mass")

fenton2 = 
  fenton %>% 
  group_by(Mass, Forest, Treatment) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup %>% 
  group_by(Treatment,Mass) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup %>% 
  group_by(Treatment, n) %>% 
  dplyr::summarise(count = n())
```


```{r fenton_nosc, fig.width=10, fig.height=5}
fenton_nosc = 
  fenton %>%
  distinct(Treatment, Mass) %>% 
  left_join(dplyr::select(meta, Mass, NOSC, Class),  by = "Mass") %>% 
  ggplot()+
  geom_histogram(aes(x = NOSC, fill = Treatment), position = "identity", alpha = 0.5)

fenton_oc = 
  fenton %>%
  distinct(Treatment, Mass) %>% 
  left_join(dplyr::select(meta, Mass, NOSC, Class, HC, OC),  by = "Mass") %>% 
  ggplot()+
  geom_histogram(aes(x = OC, fill = Treatment), position = "identity", alpha = 0.5)+
  labs(x = "O/C")


library(patchwork)

fenton_nosc + fenton_oc +
  plot_layout(guides = "collect") +
  plot_annotation(title = "oxidation states",
                  subtitle = "combined for HW and SW")
```

all lost and gained -- by forest

```{r fenton_nosc_all, fig.width=10, fig.height=5}
ggplot(loss, aes(x = NOSC, color = loss, fill = loss))+
  geom_histogram(alpha = 0.5, position = "identity")+
  facet_grid(~Forest)
```

```{r fenton_nosc_carblignin, eval=F}

loss_subset = 
  loss %>% 
  filter(Class %in% c("Carbohydrate-like","Lignin-like"))

ggplot(loss_subset)+
  geom_boxplot(aes(x = NOSC, y=80, color = loss),  width = 10)+
  geom_histogram(aes(x = NOSC, color = loss, fill = loss),
                 alpha = 0.5, position = "identity")

ggplot(loss[loss$Class=="Lignin-like",], aes(x = NOSC, color = loss, fill = loss))+
  geom_histogram(alpha = 0.5, position = "identity")+
  geom_boxplot(aes(x = NOSC, y=80, color = loss), fill = "white",  width = 10)


ggplot(fenton[fenton$Class=="Lignin-like",])+
  geom_boxplot(aes(x = OC, y=150, color = Treatment),  width = 10)+
  geom_histogram(aes(x = OC, color = Treatment, fill = Treatment),
                 alpha = 0.5, position = "identity")

```

## Goethite

```{r goethite_sorbed}
# set order of levels
sorbed2 = 
  sorbed %>% 
  group_by(Mass,fenton,adsorbed, HC, OC, Class) %>% 
  dplyr::summarise(n = n()) %>% 
  mutate(fenton = factor(fenton, levels = c("PreFenton", "PostFenton"))) %>% 
  ungroup 
  #dplyr::mutate(fenton=factor(fenton, levels = c("PreFenton","PostFenton")))

```

```{r sorbed_vk1, fig.width=9, fig.height=5}
gg_vankrev(sorbed2[sorbed2$adsorbed=="sorbed",], aes(x = OC, y = HC, color=adsorbed))+
 # scale_color_manual(values = c("#F1A340","#998EC3"))+
 scale_color_manual(values = c("darkblue"))+
  facet_wrap(~fenton)+
  theme_kp()+
  theme(legend.position = "top",
        legend.direction = "horizontal")+
  labs(title = "post-Goethite, adsorbed peaks only")+
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

```

```{r goethite_vk2, fig.width=5, fig.height=5}
# set order of levels
gg_sorbed =
  gg_vankrev(sorbed[sorbed$adsorbed=="sorbed",], aes(x = OC, y = HC, color=fenton))+
 # scale_color_manual(values = c("#F1A340","#998EC3"))+
 #scale_color_manual(values = c("grey40","darkblue"))+
  scale_color_manual(values = soil_palette("redox",2),
                     breaks = c("PreFenton", "PostFenton"))+
  #facet_wrap(~fenton)+
  stat_ellipse()+
  theme_kp()+
  theme(legend.position = c(0.65, 0.92),
        legend.direction = "horizontal")+
  guides(colour = guide_legend(override.aes = list(alpha = 1)))


ggMarginal(gg_sorbed,groupColour = TRUE,groupFill = TRUE, size = 10)

```

contribution of groups  

```{r sorbed_relabund}
sorbed_relabund %>% 
  group_by(fenton,adsorbed,Class) %>% 
  dplyr::summarise(relabund = mean(relabund)) %>% 
  ungroup %>% 
  dplyr::mutate( #fenton = factor(fenton, levels = c("PreFenton","PostFenton")), 
                  Class = factor(Class, levels = c("Condensed Ar",
                                                 "Aromatic",
                                                 "Lignin-like",
                                                 "Carbohydrate-like",
                                                 "Aliphatic-noN",
                                                 "Aliphatic+N")))->
  sorbed_relabund2
```

```{r goethite_groups, fig.width=5, fig.height=6}
ggplot(sorbed_relabund2[sorbed_relabund2$adsorbed=="sorbed",], aes(x=Class, y=relabund, fill=fenton))+
  geom_bar(stat="identity", position = position_dodge(width=0.7),width=0.5, 
           alpha=0.8,color="black",size=0.7)+
#  scale_fill_manual(values = c("#B35806","#542788"))+
#  scale_fill_manual(values = c("#F1A340","#998EC3"))+
#  scale_fill_manual(values = c("grey80","darkblue"))+
    scale_fill_manual(values = soil_palette("redox",2),
                      breaks = c("PreFenton", "PostFenton"))+

  ylab("% of total sorbed peaks")+
  xlab("")+

  theme_bw()+
  theme_kp()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, face="plain", size = 12))+
  theme(legend.position = c(0.8, 0.8))+
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

adsorbed peaks common to pre and post-Fenton

```{r sorbedpeaks_common, fig.width=6, fig.height=6}
# set order of levels
sorbed_summary = 
  sorbed %>% 
  group_by(Mass,fenton,adsorbed, HC, OC, Class) %>% 
  dplyr::summarise(n = n()) %>% 
  ungroup %>% 
  dplyr::mutate(fenton=factor(fenton, levels = c("PreFenton","PostFenton")))

sorbed_common = 
  sorbed_summary %>% 
  filter(adsorbed=="sorbed") %>% 
  group_by(Mass,fenton,adsorbed, HC, OC, Class) %>% 
  dplyr::summarise(n = n())  %>% 
  ungroup %>% 
  group_by(Mass, HC, OC) %>% 
  dplyr::mutate(common="common",
                unique = n(),
                unique2 = case_when(unique==1 ~ fenton,
                                    unique==2 ~ as.factor(common)))

gg_vankrev(sorbed_common, aes(x = OC, y = HC, color=unique2))+
  labs(title = "adsorbed peaks common to pre and post-Fenton")+
  # scale_color_manual(values = c("#F1A340","#998EC3"))+
  # scale_color_manual(values = c("grey40","darkblue"))+
  #facet_wrap(~fenton)+
  theme_kp()
```

Fenton-gained peaks vs. adsorbed

```{r postfenton_newadsorbed, fig.height=5, fig.width=5.5}
fenton_loss %>% 
  filter(loss == "gained") %>% 
  gg_vankrev(aes(x = OC, y = HC))+
  geom_point(data = 
               sorbed %>% 
               filter(fenton=="PostFenton" & adsorbed == "sorbed"), color = "orange", size=1)+
#  labs(title = "Post-Fenton adsorbed peaks",
#       subtitle = "new peaks in grey, adsorbed peaks in orange")+
  annotate("curve", x = 0.5, xend = 0.4, y = 2.3, yend = 2.15, curvature = 0.3, size=1)+
  annotate("curve", x = 0.5, xend = 0.55, y = 0.65, yend = 0.3, curvature = 0.3, color = "orange", size=1)+
  annotate("text", label = "peaks adsorbed post-Fenton", x = 0.57, y = 0.3, hjust = 0)+
  annotate("text", label = "new peaks detected post-Fenton", x = 0.52, y = 2.3, hjust = 0)+
  theme_kp()+
  NULL
```


### NOSC
calculating NOSC of adsorbed molecules

```{r}
adsorbed_nosc = sorbed %>% 
  left_join(dplyr::select(meta, Mass, NOSC),  by = "Mass")

adsorbed_nosc_subset = 
  adsorbed_nosc %>% 
  filter(Class %in% c("Carbohydrate-like","Lignin-like"),
         adsorbed %in% c("sorbed"))
#dplyr::mutate(fenton = factor(fenton, levels = c("PreFenton", "PostFenton")))


ggplot(adsorbed_nosc[adsorbed_nosc$adsorbed=="sorbed",], aes(x = NOSC, color = fenton, fill = fenton))+
  geom_histogram(alpha = 0.5, position = "identity")+
  geom_boxplot(aes(x = NOSC, y=100, color = fenton),  fill = "white", width = 10)+
  ggtitle("adsorbed")+
  facet_wrap(~Class)

adsorbed_nosc_subset %>% 
  ungroup %>% 
  distinct(Mass, fenton, Class, NOSC) %>% 
  ggplot(aes(x = NOSC, color = fenton, fill = fenton))+
  geom_histogram(alpha = 0.7, position = "identity")+
  geom_boxplot(aes(x = NOSC, y=45, color = fenton),  fill = "white", width = 5)+

  scale_fill_manual(values = soil_palette("redox",2), guide = guide_legend(reverse=TRUE))+
  scale_color_manual(values = soil_palette("redox",2), guide = guide_legend(reverse=TRUE))+
  facet_wrap(~Class)+
  theme_kp()+
  labs(title = "adsorbed peaks for carb and lignin-like molecules")
```


---


<details>
  <summary>Session Info</summary>

Run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>
